---
title: "Lab 03"
author: "Lamar Aljahdali"
date: "5/19/2023"
editor_options: 
  chunk_output_type: console
output:
  pdf_document: default
---

# Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(sf)
``` 

# Data

```{r load-data, message = FALSE}
fl_votes <- st_read("data/fl_votes.shp", quiet = TRUE)
fl_votes %>%
  slice(1:6)
```


# Exercise 1

```{r add-winner}
fl_votes_sf <- st_as_sf(fl_votes, coords = c("longitude", "latitude"), crs = 4326)

fl_votes_sf <- fl_votes_sf %>% 
  mutate(winner20 = if_else(dem20 > rep20, "Democrat",
                            if_else(dem20 < rep20, "Republican", "Tie")))

```


# Exercise 2

```{r fl-plot-1}
ggplot() +
  geom_sf(data = fl_votes_sf, aes(fill = winner20)) +
  scale_fill_manual(values = c("Democrat" = "#0015BC", "Republican" = "#DE0100", "Tie" = "gray"),
                    na.value = "gray") +
  labs(fill = "Winner 2020") +
  theme_minimal()

```

# Exercise #3

```{r fl-props}
fl_votes_sf <- fl_votes_sf %>% 
  mutate(prop_rep16 = rep16 / (rep16 + dem16),
         prop_rep20 = rep20 / (rep20 + dem20))
```

# Exercise 4

```{r fl-plot-2}
ggplot() +
  geom_sf(data = fl_votes_sf, aes(fill = prop_rep20)) +
  scale_fill_gradient(low = "#DE0100", high = "#0015BC",
                      na.value = "gray") +
  labs(fill = "Republican Proportion 2020") +
  theme_minimal()
```

# Exercise 5

```{r fl-change}
fl_votes_sf <- fl_votes_sf %>% 
  mutate(diff_rep = prop_rep20 - prop_rep16)

```

# Exercise 6

```{r fl-plot-3}
ggplot() +
  geom_sf(data = fl_votes_sf, aes(fill = diff_rep)) +
  scale_fill_gradient2(low = "#0015BC", mid = "gray", high = "#DE0100",
                       midpoint = 0, na.value = "gray",
                       guide = guide_colorbar(title = "Vote Share Change")) +
  labs(fill = "Vote Share Change") +
  theme_minimal()
```

# Exercise 7
The visualizations developed provide insights into the 2016 and 2020 Presidential elections in Florida. 
The first plot (Exercise 2) shows the winner of each county in the 2020 election, with Democrat, Republican, and Tie categories represented by different colors. 
The second plot (Exercise 4) displays the proportion of the two-party vote cast for the Republican candidate in 2020, with a gradient color scheme from red to blue. 
The third plot (Exercise 6) depicts the change in Republican vote share between 2016 and 2020, with a color gradient from blue to red.

Limitations of these visualizations include:
1. The visualizations focus only on Republican and Democratic candidates, without considering other parties or independent candidates.
2. The color schemes chosen for the plots are subjective and may not be universally interpretable.
3. The plots do not take into account the population differences between counties, which can impact the overall representation of results.



